Computing with cut systems +

### 2015-10-14 cutsystems-scheme.tex
Revisions.

\documentclass[english,letter paper,12pt,leqno]{article}
\usepackage{stmaryrd}
\usepackage{amsmath, amscd, amssymb, mathrsfs, accents, amsfonts,amsthm}
\usepackage[all]{xy}
\usepackage{tikz}
\def\nicedashedcolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=gray, dashed]}
\def\nicecolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=white]}
\def\nicepalecolourscheme{\shadedraw[top color=blue!12, bottom color=blue!12, draw=white]}
\def\nicenocolourscheme{\shadedraw[top color=gray!2, bottom color=gray!25, draw=white]}
\def\nicereallynocolourscheme{\shadedraw[top color=white!2, bottom color=white!25, draw=white]}
\definecolor{Myblue}{rgb}{0,0,0.6}
\usepackage[a4paper,colorlinks,citecolor=Myblue,linkcolor=Myblue,urlcolor=Myblue,pdfpagemode=None]{hyperref}

@@

\section{Introduction}

In this paper we develop a new semantics of composition using a bicategorical formalism that we call \emph{cut systems}, inspired by Gentzen's cut-elimination \cite{gentzen} and Girard's geometry of interaction \cite{girard_towards}. The formalism is used to ``make computable'' a bicategory of isolated hypersurface singularities and matrix factorisations that has been studied in the context of topological field theory \cite{McNameethesis, cr0909.4381}. This means that the operations making up the data of the bicategory are given algorithmically, and can be implemented on a computer. One of the main applications, to be given in a sequel \cite{cutsystems2}, is to construct a model of a fragment of Girard's linear logic \cite{girard_llogic}. Roughly speaking, this provides mathematical semantics for a programming language in which data types are interpreted as hypersurface singularities, and programs are interpreted as homotopy equivalence classes of matrix factorisations.

What is a semantics of composition? Consider that in a category or bicategory $\cat{A}$, for every triple of objects $a,b,c$ and morphisms
\[
\xymatrix@C+2pc{
a \ar[r]^-{\Phi} & b \ar[r]^-{\Psi} & c
}
\]
there is an assignment to $(\Psi, \Phi)$ of a composition $\Psi \circ \Phi$. A \emph{semantics of composition} for $\cat{A}$ is a mathematical construct providing some kind of inner structure for this ``assignment'' which is coherent as $a,b,c,\Phi,\Psi$ vary. The main examples of this kind of semantics are in computer science and mathematical logic.

For example, suppose that $\cat{A}$ is the category of sets and that we have functions
\[
\xymatrix@C+2pc{
\mathbb{N} \ar[r]^-{\Phi} & \mathbb{N} \ar[r]^-{\Psi} & \mathbb{N}
}
\]
where $\Phi$ and $\Psi$ are ``finitely described'' in the sense that they are computed by algorithms $X$ and $Y$, respectively. To be explicit, let us take $X,Y$ to be terms in the $\lambda$-calculus \cite{selinger}. If $[-]$ denotes assignment to an algorithm of the function it describes, then $\Phi = [X], \Psi = [Y]$ and $\Psi \circ \Phi = [Y \l X]$ where $Y \l X$ is the result of feeding the output of the algorithm $X$ as an input to the algorithm $Y$. In the logic literature this concatenation is known as a \emph{cut}.\footnote{We note that the notation $Y \l X$ for the cut is not standard, so don't blame the logicians for it.} To encounter the desired semantics we need the additional observation that the cut $Y \l X$ is not in normal form (whatever this means) but it may be reduced to normal form by a rewriting process on terms in the $\lambda$-calculus called $\beta$-reduction. This does not change the associated function: the normal form of $Y \l X$ also computes $\Psi \circ \Phi$.

A semantics of composition for the category of computable functions is then given by the process which begins with the pair $(Y,X)$ of finite descriptions of $\Psi,\Phi$, forms the cut $Y \l X$, and then computes a normal form giving a finite description of $\Psi \circ \Phi$.

\medskip

This seems very far from geometry, but we can informally recast a problem of algebraic geometry in a similar light. Suppose $A,B,C$ are smooth projective algebraic varieties and that we are given triangulated functors between derived categories of coherent sheaves
\[
\xymatrix{
\mathbf{D}^b(A) \ar[r]^-{\Phi} & \mathbf{D}^b(B) \ar[r]^-{\Psi} & \mathbf{D}^b(C)
}\,.
\]
Suppose that $\Phi,\Psi$ are finitely described, in the sense that they are both integral functors associated to kernels: $\Phi$ is convolution with a complex of coherent sheaves $X$ on $A \times B$, and $\Psi$ is convolution with a complex $Y$ on $B \times C$. If for example $A,B,C$ are the $\operatorname{Proj}$ of graded algebras, $X$ and $Y$ can be described by lists of matrices of homogeneous polynomials. An obvious question is: what is the finite description of the functor $\Psi \circ \Phi$?

It is easy to see that $\Psi \circ \Phi$ is convolution with the kernel $Y \otimes^{\mathbb{L}}_{\cat{O}_B} X$, which is a complex of quasi-coherent sheaves on $A \times C$ with coherent cohomology, but this is not described by a finite amount of data. Finding a finite description of $\Psi \circ \Phi$ is precisely the problem of determing the complex of coherent sheaves which is isomorphic, in $\mathbf{D}^b(A \times C)$, to $Y \otimes^{\mathbb{L}}_{\cat{O}_B} X$. We refer to such a complex of coherent sheaves as a \emph{finite model} of $Y \otimes^{\mathbb{L}}_{\cat{O}_B} X$.

Consider the case where $A = C = \operatorname{Spec}(\mathbb{C})$, so $X,Y$ are complexes of coherent sheaves on $B$ and $Y \otimes^{\mathbb{L}}_{\cat{O}_B} X \cong \mathbb{R}\! \operatorname{Hom}_{\cat{O}_B}(Y^{\vee}, X)$ is a complex of infinite-dimensional vector spaces with finite-dimensional cohomology. An example of a finite model is an isomorphism in the derived category of all $\mathbb{C}$-vector spaces of this complex with its cohomology
\begin{equation}\label{eq:computing_hom_derived}
Y \otimes^{\mathbb{L}}_{\cat{O}_B} X \cong \bigoplus_{n \in \mathbb{Z}} \Ext^n(Y^{\vee},X)[-n]\,.
\end{equation}
One finite description of $Y \otimes^{\mathbb{L}}_{\cat{O}_B} X$, and transitively of $\Psi \circ \Phi$, is therefore a list of numbers: the dimensions of the vector spaces on the right hand side of \eqref{eq:computing_hom_derived}. Determining these numbers is difficult, to put it mildly: a fundamental problem in algebraic geometry is to understand how these numbers depend on the input data $X$ and $Y$.

By a (perhaps naive) analogy with computable functions, one might wonder if in the general case there exists an object $Y \l X$ in some triangulated category together with a reduction process that would help find a finite description for $\Psi \circ \Phi$. This is the question of whether or not there is a semantics of composition in the bicategory of smooth projective varieties and Fourier-Mukai kernels \cite{ct1007.2679}. Unfortunately, we do not know such a theory!

However, it turns out that there \emph{is} a satisfactory semantics of this kind when instead of smooth projective varieties we consider isolated hypersurface singularities $W,V,U$ and triangulated functors between homotopy categories of matrix factorisations \cite{yoshino98}
\begin{equation}\label{eq:pair_functors_hmf}
\xymatrix{
\hmf(W) \ar[r]^-{\Phi} & \hmf(V) \ar[r]^-{\Psi} & \hmf(U)
}\,.
\end{equation}
To explain further, we need to introduce the language of cut systems.

\medskip

A \emph{cut system} $\cat{A}$ over a $\mathbb{Q}$-algebra $k$ has a class of objects and for each pair $a,b$ of objects a $k$-linear supercategory $\cat{A}(a,b)$. Informally, one should think of this as a category of finite models. There is an operation on $\cat{A}$, the \emph{cut}, which assigns to $X \in \cat{A}(a,b)$ and $Y \in \cat{A}(b,c)$ a representation $Y \l X$ in $\cat{A}(a,c)$ of a Clifford algebra which depends on the intermediate object $b$. The object $Y \l X$ and the endomorphisms giving the Clifford action are required to be effectively computable, say by explicit formulas, from the input data $X,Y$. In the example of a cut system developed in this paper, Clifford algebras arise as endomorphism algebras of exterior algebras, which in turn arise as the underlying graded modules of Koszul complexes. Associated to the cut system $\cat{A}$ is a bicategory $\cat{A}^\bullet$ in which the composition is derived from the cut operation, and so is also effectively computable.

A \emph{cut model} of a bicategory $\cat{B}$ is a cut system $\cat{A}$ with the same objects as $\cat{B}$, together with an equivalence $I: \cat{B} \lto \cat{A}^\bullet$. There are fully faithful functors $I: \cat{B}(a,b) \lto \cat{A}(a,b)$ which we think of as assigning finite models to the $1$-morphisms of $\cat{B}$, and an isomorphism of Clifford representations for every composable pair of $1$-morphisms $X,Y$ in $\cat{B}$, where $S$ denotes the spinor representation:
\begin{equation}\label{eq:intro_cut_semantics}
I(Y) \l I(X) \cong S \otimes_k I( Y \circ X )\,.
\end{equation}
The cut model provides $\cat{B}$ with a semantics of composition: given finite models $I(X), I(Y)$ of $X,Y$ there is a process beginning with the cut $I(Y) \l I(X)$ which extracts the subobject $I(Y \circ X)$ using the Clifford action. The result $I( Y \circ X )$ of this computation is the desired finite model of the composition $Y \circ X$ in $\cat{B}$.
\medskip

The formalism is adapted to, and motivated by, an example: the bicategory of Landau-Ginzburg models $\LG$. Let us return to the diagram \eqref{eq:pair_functors_hmf} as our starting point. There is an analogue of integral functors in this setting, so that it makes sense to talk about $\Phi$ and $\Psi$ as being respectively given by convolution with kernels $X,Y$. If $W \in k[x_1,\ldots,x_n]$ and $V \in k[z_1,\ldots,z_m]$ define isolated hypersurface singularities then $X$ is a finite rank matrix factorisation of $V - W$ over $k[x,z]$, and similarly for $Y$.

Again, the question is: what is the finite description of $\Psi \circ \Phi$? It is naturally given by convolution with a matrix factorisation $Y \otimes X$ which is not finitely generated: finding a finite description of $\Psi \circ \Phi$ therefore amounts to finding a finite model for $Y \otimes X$. To fit this into the formalism of cut systems, we use the  bicategory $\LG$ in which the objects are isolated hypersurface singularities\footnote{Roughly speaking. See Section \ref{section:superbicatLG}.} and $1$-morphisms $W \lto V$ are matrix factorisations of $V - W$. In this bicategory, the finite descriptions of the functors in \eqref{eq:pair_functors_hmf} give rise to a composable pair of $1$-morphisms
\begin{equation}\label{eq:cut_lg}
\xymatrix@C+2pc{
W \ar[r]^-{X} & V \ar[r]^-{Y} & U
}\,.
\end{equation}
whose composition in $\LG$ is the tensor product $Y \otimes X$.

The main results of this paper are the construction of a cut sytem $\cat{L}$ and cut model $I: \LG \lto \cat{L}^\bullet$. In the situation of \eqref{eq:cut_lg} $X$ and $Y$ are their own finite model (so $I(X) = X$ and $I(Y) = Y$) and a finite model of the composite $Y \otimes X$ in $\cat{B}$ may be computed from the cut $Y \l X$ in $\cat{L}$, which is an explicitly given finite rank matrix factorisation of $U - W$ together with a Clifford action described in terms of derivatives of the differentials of $X,Y$ and certain invariants called Atiyah classes.

One consequence is that $\LG$ is equivalent to the bicategory $\cat{L}^\bullet$, which is \emph{computable} in the sense that its operations may be embodied in software. To some extent this software already exists: a component of what we are here calling the cut system $\cat{L}$ (developed earlier in joint work with Dyckerhoff \cite{dm1102.2957}\footnote{For a more detailed discussion of the relation of this work with the present paper, see Remark \ref{remark:relation_to_toby_paper}.}) has already been implemented in the computer algebra package Singular in joint work with Carqueville \cite{khovhompaper}, and used to compute tables of Khovanov-Rozansky knot homologies. The code was also used in the discovery of the generalised orbifolding relations among ADE singularities by Carqueville, Ros Camacho and Runkel \cite{ade}.

\medskip
In conclusion, the purpose of this paper is to develop the theory of cut systems as a mathematical foundation for the application to be given in \cite{cutsystems2}. Let us briefly summarise the contents of that work: one can show that the superbicategory $\cat{L}^\bullet$ is monoidal, has adjoints and is pivotal. This makes it a natural target for models of Girard's linear logic, which is a generalisation of the $\lambda$-calculus in which the terms may be viewed as algorithms that have a refined syntax of resource usage compared to the $\lambda$-calculus. The model of linear logic to be given in \cite{cutsystems2} interprets algorithms $x,y$ as $1$-morphisms $[x],[y]$ in $\cat{L}^\bullet$.

This interpretation has two novelties: first of all, the target $\cat{L}^\bullet$ has a semantics of composition that may be compared to cut-elimination in linear logic (the analogue of $\beta$-reduction in the $\lambda$-calculus) as in Girard's geometry of interaction \cite{girard_towards}. This connection is what motivates the terminology of ``cut systems'' in the present paper. The second novelty is that, compared to the usual mathematical semantics of programming languages where algorithms are interpreted as functions between sets, our interpretation is bicategorical: in particular, equations $[x] \circ [y] \cong [y] \circ [x]$ or $[x]^2 \cong [x]$ on programs are represented by data that is part of the semantics.

%. For a precise definition see Section \ref{??}}. To such a $W$ we associate a category of algebraic objects called matrix factorisations. A matrix factorisation of $W(x)$ is a matrix of polynomials $D$ satisfying $D^2 = W \cdot I_n$, where $I_n$ is the identity matrix.

%For $W = x y^2 + y^3$ an example is
%\[
%D_1 = \begin{pmatrix} 0 & x - i y \\ x + iy & 0 \end{pmatrix} %, \qquad D_2 = \begin{pmatrix} 0 & y - x \\ (y - e^{2 \pi i /3} x)(y - e^{4 \pi i /3} x ) & 0 \end{pmatrix}
%\]
%A \emph{morphism} of matrix factorisations $\Phi: D_1 \lto D_2$ is a matrix of polynomials $\Phi$ satisfying the relation $\Phi D_1 = D_2 \Phi$. Together the matrix factorisations of $W$ and the morphisms between them form the category of matrix factorisations of $W$.

%In the second part of the paper we explain how the multiplicative fragment of linear logic can be interpreted in the bicategory of Landau-Ginzburg models, with formulas mapped to isolated hypersurface singularities and proofs mapped to matrix factorisations. The multiplicative fragment is not very expressive, so our focus in the present paper is not on embedding sophisticated programs into $\cat{LG}$. Rather, we explain how cut systems give a dynamic model of cut-elimination following Girard's geometry of interaction. The basic construction of a cut system involves homological perturbation, and by interpreting this as a fixed point construction we will show that the result of normalization can be expressed in a power series expansion similar to Girard's famous execution formula.

\medskip

The paper is structured as follows: in Section \ref{section:background} we recall the definition of superbicategories and Clifford algebras, and introduce the Clifford thickening of a supercategory. In Section \ref{section:cut_systems} we define cut systems and the superbicategory associated to a cut system, then we construct the cut model of $\LG$ in Section \ref{section:lg_cut_system}.

\medskip

\emph{Acknowledgements.} Thanks to Nils Carqueville and Ingo Runkel for helpful comments on the draft. Although the sequel on linear logic is the proper place to fully acknowledge the debt, we cannot neglect to mention the important role that the beautiful ideas in Girard's geometry of interaction program \cite{girard_towards} had in shaping the present work.

\section{Background}\label{section:background}

Throughout $k$ is a commutative $\mathbb{Q}$-algebra. By default categories and functors are $k$-linear.

\subsection{Supercategories}

A superbicategory has for every $1$-morphism $F$ an associated $1$-morphism $\Psi F$ with $\Psi^2 \cong 1$, and in this sense the categories of $1$-morphisms are $\mathbb{Z}_2$-graded. The bicategory of Landau-Ginzburg models is an example of a superbicategory. Since Clifford algebras (which are $\mathbb{Z}_2$-graded) play a fundamental role in the definition of cut systems, superbicategories are the natural language. Our main references for the foundations are \cite{ellis_lauda,kang,kang2}, but we give the full definitions below as these references work only with strict bicategories.

\begin{definition} A \emph{supercategory} is a category $\cat{C}$ together with a functor $\Psi: \cat{C} \lto \cat{C}$ and a natural isomorphism $\xi: \Psi^2 \lto 1_{\cat{C}}$ satisfying the condition
\[
\xi * 1_{\Psi} = 1_{\Psi} * \xi
\]
as natural transformations $\Psi^3 \lto \Psi$. A \emph{superfunctor} $(F, \gamma)$ from a supercategory $(\cat{C}, \Psi_{\cat{C}})$ to a supercategory $(\cat{D}, \Psi_{\cat{D}})$ is a functor $F: \cat{C} \lto \cat{D}$ together with a natural isomorphism $\gamma: F \Psi_{\cat{C}} \lto \Psi_{\cat{D}} F$ satisfying
\[
1_F * \xi = (\xi * 1_F ) ( 1_{\Psi} * \gamma ) ( \gamma * 1_{\Psi} )\,.
\]
A \emph{supernatural transformation} $\varphi: F \lto G$ between superfunctors $(F,\gamma_F), (G,\gamma_G)$ is a natural transformation making the following diagram commute:
\[
\xymatrix{
F \Psi \ar[r]^-{\varphi \Psi}\ar[d]_-{\gamma_F} & G \Psi \ar[d]^-{\gamma_G}\\
\Psi F \ar[r]_-{\Psi \varphi} & \Psi G
}\,.
\] 
\end{definition}

\begin{example}\label{example:Aassup} If $A$ is a $\mathbb{Z}_2$-graded $k$-algebra there is a supercategory $\cat{A}$ with set of objects $|\cat{A}| = \mathbb{Z}_2$ and morphisms $\cat{A}(0,0) = \cat{A}(1,1) = A_0, \cat{A}(0,1) = \cat{A}(1,0) = A_1$ with the obvious composition and addition rules. The functor $\Psi$ is defined on objects by $\Psi(i) = i+1$ and on morphisms by the identity, while $\xi = 1$.
\end{example}

Let $\cat{C}$ be a supercategory. For objects $X,Y \in \cat{C}$ we define the $\mathbb{Z}_2$-graded $k$-module
\[
\cat{C}^*(X,Y) = \prod_{i \in \mathbb{Z}_2} \cat{C}(X, \Psi^i Y)\,.
\]
There is for any triple $X,Y,Z$ of objects a morphism of graded modules
\[
\cat{C}^*(Y,Z) \otimes_k \cat{C}^*(X, Y) \lto \cat{C}^*(X,Z)
\]
given for instance by $\varphi \otimes \psi \mapsto \xi \Psi( \varphi ) \psi$ when $\varphi, \psi$ are both of degree one. This makes $\cat{C}$ into a category enriched over the monoidal category of $\mathbb{Z}_2$-graded $k$-modules.

There are isomorphisms of graded modules $\cat{C}^*(X, \Psi Y) \cong \Psi \cat{C}^*(X, Y) \cong \cat{C}^*(\Psi X, Y)$ where $\Psi V = V [1]$ denotes the grading shift on a $\mathbb{Z}_2$-graded $k$-module. For $\mathbb{Z}_2$-graded $k$-modules $V,W$ we write $\Hom^0_k(V, W)$ for the module of degree zero maps and $\Hom_k^*(V, W)$ for the $\mathbb{Z}_2$-graded module of homogeneous maps. If $A$ is a $\mathbb{Z}_2$-graded $k$-algebra then $A^{\operatorname{op}}$ denotes the same underlying graded module with multiplication $f * g = (-1)^{|f||g|} gf$.

\begin{definition} Let $A$ be a $\mathbb{Z}_2$-graded $k$-algebra and $\cat{C}$ a supercategory. A \emph{left $A$-module in $\cat{C}$} is an object $X \in \cat{C}$ together with a morphism of $\mathbb{Z}_2$-graded algebras $A \lto \cat{C}^*(X,X)$. A \emph{morphism} of left $A$-modules is a morphism in $\cat{C}$ which commutes with the $A$-action in the obvious sense. The category of left $A$-modules in $\cat{C}$ is denoted $\cat{C}_A$. Right modules are defined similarly, using $A^{\operatorname{op}}$.
\end{definition}

\begin{remark}\label{remark:supercat_idempcomp} Viewing $A$ as a supercategory $\cat{A}$ as in Example \ref{example:Aassup}, there is an equivalence between the category of $A$-modules in $\cat{C}$ and the category of superfunctors $\cat{A} \lto \cat{C}$ and supernatural transformations.
\end{remark}

\begin{remark}\label{remark:idempotent_completion} The \emph{idempotent completion} $\cat{C}^\omega$ of a category $\cat{C}$ has as objects pairs $(X,e)$ consisting of an object $X \in \cat{C}$ and an idempotent endomorphism $e: X \lto X$. In $\cat{C}^\omega$ a morphism $\phi: (X,e) \lto (Y,e)$ is a morphism $\phi: X \lto Y$ satisfying $\phi e = \phi$ and $e \phi = \phi$. The identity on $(X,e)$ is $e$. There is a fully faithful functor $\iota: \cat{C} \lto \cat{C}^{\omega}, \iota(X) = (X,1_X)$ which is an equivalence if $\cat{C}$ is idempotent complete (i.e. all idempotents split in $\cat{C}$).

If $\cat{C}$ is a supercategory then there is a functor $\Psi^\omega: \cat{C}^\omega \lto \cat{C}^\omega$ defined on objects by $\Psi(X,e) = (\Psi X, \Psi e)$ and a natural isomorphism $\xi: \Psi^\omega \circ \Psi^\omega \lto 1_{\cat{C}^\omega}$ defined by $\xi_{(X,e)} = e \xi_X$. The tuple $(\cat{C}^\omega, \Psi^\omega, \xi^\omega)$ is a supercategory.
\end{remark}

For background on bicategories see \cite{bor94} and the references in \cite{lgdual}. We follow the notation of \cite{lgdual}, so that lower case letters $a,b, \ldots$ denote objects of a bicategory, while upper case letters $X,Y, \ldots$ and greek letters $\alpha, \beta, \ldots$ respectively denote $1$-morphisms and $2$-morphisms. Units are denoted $\Delta$, associators are $\alpha$, and unitors are $\lambda, \rho$. The composition of $Y,X$ is denoted $YX$ or $Y \circ X$.

\begin{definition} A \emph{superbicategory} is a bicategory $\cat{B}$ together with the data:
\begin{itemize}
\item For each object $a$ a $1$-morphism $\Psi_a: a \lto a$ and a $2$-isomorphism $\xi_a: \Psi^2_a \lto \Delta_a$.
\item For each $1$-morphism $X: a \lto b$ a natural $2$-isomorphism $\gamma_X: X \Psi_a \lto \Psi_b X$.
\end{itemize}
This data is required to satisfy the following axioms:
\begin{itemize}
\item For each composable pair $X,Y$ of $1$-morphisms the diagram
\[
\xymatrix@C+2pc@R+1pc{
(YX) \Psi \ar[rrr]^-{\gamma_{YX}} \ar[d]_-{\alpha} & & & \Psi ( YX )\\
Y ( X \Psi ) \ar[r]_-{1_Y * \gamma_X} & Y( \Psi X ) \ar[r]_-{\alpha^{-1}} & ( Y \Psi ) X \ar[r]_-{\gamma_Y * 1_X} & (\Psi Y ) X \ar[u]_-{\alpha}
}
\]
commutes.
\item For every object $a$, $\xi_a * 1_\Psi = 1_\Psi * \xi_a$.
\item For every $1$-morphism $X : a \lto b$, $1_X * \xi_a = ( \xi_b * 1_X ) ( 1_\Psi * \gamma_X ) (\gamma_X * 1_\Psi )$.
\end{itemize}
\end{definition}

\begin{example} Supercategories, superfunctors and supernatural transformations form a superbicategory $\Cat^{\operatorname{sup}}_k$.
\end{example}

A superbicategory can be constructed out of a bicategory $\cat{B}$ in which the categories $\cat{B}(a,b)$ are all equipped with the structure of supercategories; see Appendix \ref{section:constructing_superbicategories}.

\begin{example}\label{example:bicategory_m} There is a bicategory of $\mathbb{Z}_2$-graded $k$-algebras where $1$-morphisms are $\mathbb{Z}_2$-graded bimodules and $2$-morphisms are degree zero bimodule maps. Given a $B$-$A$-module $M$ the shift $M[1]$ has the grading $M[1]_i = M_{i+1}$ and the left and right action given by
\[
b \cdot m = (-1)^{|b|} bm, \qquad m \cdot a = (-1)^{|a|} m a\,.
\]
This is a functor $\Psi = (-)[1]$ on the category of $B$-$A$-bimodules and with $\xi = 1$ this defines the structure of a supercategory on this category of bimodules. The usual isomorphisms of bimodules $\tau$
\begin{align}
N[1] \otimes M &\lto (N \otimes M)[1], \qquad n \otimes m \mapsto n \otimes m \label{eq:shift_iso_a}\\
N \otimes M[1] &\lto (N \otimes M)[1], \qquad n \otimes m \mapsto (-1)^{|n|} n \otimes m\label{eq:shift_iso_b}
\end{align}
satisfy the conditions in Appendix \ref{section:constructing_superbicategories} and therefore give the bicategory of $\mathbb{Z}_2$-graded algebras and bimodules the structure of a superbicategory.
\end{example}

>>>>

\documentclass[english,letter paper,12pt,leqno]{article}
\usepackage{stmaryrd}
\usepackage{amsmath, amscd, amssymb, mathrsfs, accents, amsfonts,amsthm}
\usepackage[all]{xy}
\usepackage{dsfont}
\usepackage{tikz}
\def\nicedashedcolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=gray, dashed]}
\def\nicecolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=white]}
\def\nicepalecolourscheme{\shadedraw[top color=blue!12, bottom color=blue!12, draw=white]}
\def\nicenocolourscheme{\shadedraw[top color=gray!2, bottom color=gray!25, draw=white]}
\def\nicereallynocolourscheme{\shadedraw[top color=white!2, bottom color=white!25, draw=white]}
\definecolor{Myblue}{rgb}{0,0,0.6}
\usepackage[a4paper,colorlinks,citecolor=Myblue,linkcolor=Myblue,urlcolor=Myblue,pdfpagemode=None]{hyperref}

@@

\tableofcontents

\section{Composition as a morphism of schemes}

We have now shown that $\LG$ is equivalent to $\cat{L}$, and in this section we explain how to model the composition of $1$-morphisms in this latter bicategory as a morphism of schemes. Throughout $k$ is a $\nQ$-algebra and $R$ any $k$-algebra.

\begin{definition}\label{defn:hmfomegac} Given $W \in R$ the category $\cat{K}(R,W)$ has for its objects tuples
\be\label{eq:hmfomegacd}
\left( D, E, l, m, \{ \gamma_i, \gamma_i^{\dagger} \}_{i=1}^m, \{ s_{ij}, s^{\dagger}_{ij}, t_{ij} \}_{i,j=1}^m, \{ z_{ai}, z^\dagger_{ai} \}_{0 \le a \le 1, 1 \le i \le m} \right)
\ee
where $m \ge 0$ is an integer and
\begin{itemize}
\item A matrix $D \in M_{2r}(R)$ with $D^2 = W \cdot I_{2r}$.
\item A morphism $E: D \lto D$
\item An odd matrix $l$ with
\be\label{eq:defnhmfomega2}
EE - E = Dl + lD\,.
\ee
\item $\gamma_i, \gamma_i^{\dagger} \in M_{2r}(R)$ are closed odd matrices, i.e. morphisms $D \lto D[1]$ for $1 \le i \le m$.
\item The $s_{ij}, s^\dagger_{ij}, t_{ij} \in M_{2r}(R)$ are odd matrices satisfying, for $1 \le i,j \le m$,
\begin{align}
\gamma_i \gamma_j + \gamma_j \gamma_i &= D s_{ij} + s_{ij} D \,,\label{eq:hmfomegac3}\\
\gamma^\dagger_i \gamma^\dagger_j + \gamma^\dagger_j \gamma^\dagger_i &= D s^\dagger_{ij} + s^\dagger_{ij} D\,,\label{eq:hmfomegac4}\\
\gamma_i \gamma_j^\dagger + \gamma_j^\dagger \gamma_i &= \delta_{ij} \cdot I_{2r} + D t_{ij} + t_{ij} D\,.\label{eq:hmfomegac5}
\end{align}
These homotopies establish that the $\gamma_i, \gamma^\dagger_i$ satisfy Clifford relations.
\item The $z_{ai}, z^\dagger_{ai} \in M_{2r}(R)$ are odd matrices satisfying, for $1 \le i \le m$,
\begin{align}
\gamma_i E - \gamma_i &= D z_{0i} + z_{0i} D\,\label{eq:hmfomegac6}\\
E \gamma_i - \gamma_i &= D z_{1i} + z_{1i} D\,\label{eq:hmfomegac7}\\
\gamma^\dagger E - \gamma^\dagger_i &= Dz^\dagger_{0i} + z^\dagger_{0i} D\,\label{eq:hmfomegac8}\\
E \gamma^\dagger - \gamma^\dagger_i &= Dz^\dagger_{1i} + z^\dagger_{1i} D\,.\label{eq:hmfomegac9}
\end{align}
These homotopies establish that the $\gamma_i, \gamma^\dagger_i$ are morphisms in $\hmf(R,W)^\omega$.
\end{itemize}
A \emph{morphism} of such tuples
\[
\phi: \left( D,E,l, \{ \gamma_i, \gamma_i^{\dagger} \}_{i=1}^m, \ldots \right) \lto \left( D',E',l', \{ \rho_i, \rho_i^{\dagger} \}_{i=1}^m, \ldots \right)
\]
is a morphism $\phi: D \lto D'$ of matrix factorisations with the additional property that $E' \phi \simeq \phi$ and $\phi E \simeq E$ and there are homotopies
\be
\phi \gamma_i \simeq 0\,, \qquad \rho_i^\dagger \phi \simeq 0
\ee
These morphisms are composed by matrix multiplication.
\end{definition}

\begin{proposition} For potentials $(x,W),(y,V)$ there is an equivalence of categories
\be\label{eq:equivalence_psi_final}
\Psi: \cat{K}( k[x,y], V - W ) \lto \cat{L}( W, V )\,.
\ee
\end{proposition}

\begin{definition} Given a $k$-algebra $R$ and integers $r \ge 1,m \ge 0, d \ge 0$ let
\[
\mathfrak{D}(R, k[x_1,\ldots,x_n],W)_{r,m,d}
\]
denote the set of tuples \eqref{eq:hmfomegacd} over the ring $R[x]$ with all matrices of fixed size $2r$ and such that every matrix in the tuple has entries which are polynomials in $R[x]$ which involve only monomials $x^\alpha$ with $|\alpha| \le d$. Here $m$ is the number of Clifford actions.
\end{definition}

\begin{lemma} Given $W \in k[x_1,\ldots,x_n]$ and integers $r \ge 1, m \ge 0, k \ge 0$ there is a finitely generated $k$-algebra $M(k[x],W)_{r,m,d}$ together with a natural bijection for any $k$-algebra $R$
\[
\Hom_{k}( M(k[x],W)_{r,m,d}, R ) \cong \mathfrak{D}( R, k[x], W )_{r,m,d}\,.
\]
\end{lemma}

\begin{definition} Given $W \in k[x_1,\ldots,x_n]$ and $r \ge 1,m \ge 0, d \ge 0$ we define the affine $k$-scheme
\[
\cat{M}( k[x], W )_{r,m,d} = \Spec\big( M(k[x],W)_{r,m, d} \big)\,.
\]
By definition there is an identification of the $k$-points of this scheme with a subset of the objects of the category $\cat{K}( k[x], W )$ since
\begin{align*}
\Big( \cat{M}( k[x], W )_{r,m,d} \Big)_k &\cong \Hom_k( M(k[x],W)_{r,m,d}, k )\\
&\cong \mathfrak{D}( k, k[x], W )_{r,m,d}\\
&\subseteq \cat{K}( k[x], W )\,.
\end{align*}
\end{definition}

\begin{theorem} Given potentials $(x,W), (y,V), (z,U)$ and $r,r' \ge 1$ and $m,m',d,d' \ge 0$ set
\begin{align*}
\cat{M}^{W,V}_{r,m,d} &= \cat{M}( k[x,y], V - W )_{r,m,d}\,,\\
\cat{M}^{V,U}_{r',m',d'} &= \cat{M}( k[y,z], U - V )_{r',m',d'}\,,\\
\cat{M}^{U,W}_{a, b,c} &= \cat{M}( k[x,z], U - W )_{a,b,c}\,.
\end{align*}
There are functions $f,g,h: \mathbb{N}^2 \lto \mathbb{N}$ and morphisms of affine schemes
\[
C_{r,r',m,m',d,d'}: \cat{M}^{V,U}_{r,d} \times \cat{M}_{r',d'}^{W,V} \lto \cat{M}_{f(r,r'), g(m,m'),h(d,d')}^{U, W}
\]
such that the diagram of sets
\[
\xymatrix@C+1pc{
\big( \cat{M}^{V,U}_{r,m,d} \big)_k \times \big( \cat{M}_{r',m',d'}^{W,V} \big)_k \ar[d] \ar[r]^{C_{r,r',m,m',d,d'}} & \big( \cat{M}_{f(r,r'), g(m,m'),h(d,d')}^{U, W} \big)_k \ar[d]\\
\cat{K}( k[y,z], U - V ) \times \cat{K}( k[x,y], V - W ) \ar[d]_-{\Psi \times \Psi} & \cat{K}( k[x,z], U - W )\ar[d]^-{\Psi} \\
\cat{L}( V, U ) \times \cat{L}( W, V ) \ar[r]_-{-\l-} & \cat{L}(W,U)
}
\]
commutes.
\end{theorem}

\begin{itemize}
\item The morphisms of schemes $C$ satisfy associativity on the nose. 
\item Given $(x,W)$ there are integers $r,d$ depending on $(x,W)$ and a morphism of schemes
\[
\Spec(k) \lto \cat{M}^{W,W}_{r,d}
\]
which embeds the unit $1$-morphism, but we don't know any reasonable way to impose a condition involving this.
\end{itemize}

Set $\cat{M} = \cat{M}^{W,V}$ and let $\mathscr{H}\!om$ denote the Clifford bundle on $\cat{M} \times \cat{M}$ whose fiber over $( [X], [X'] )$ is $X^{\vee} \l X'$. Consider
\[
\xymatrix@C+3pc{
\cat{M} \times \cat{M} \times \cat{M} \ar[r]^-{1 \times \Delta \times 1} & \cat{M} \times \cat{M} \times \cat{M} \times \cat{M}
}
\]
on the right hand side we have the vector bundle $\mathscr{H}\!om \boxtimes \mathscr{H}\!om$ which has an action of the bundle of Clifford algebras $C_m \boxtimes C_m$. Set $\varphi = 1 \times \Delta \times 1$ and consider the morphism of bundles of Clifford algebras given by the product
\be\label{eq:morphism_of_clifford}
\varphi^*( C_m \boxtimes C_m ) \lto C_m
\ee
Let $\mathscr{H}\!om$ be the bundle on $\cat{M}^3$ whose fiber over $([X_1], [X_2], [X_3])$ is $X_1^{\vee} \l X_3$. There is a morphism of vector bundles on $\cat{M}^3$
\[
\varphi^*\big( \mathscr{H}\!om \boxtimes \mathscr{H}\!om \big) \lto \mathscr{H}\!om
\]
compatible with \eqref{eq:morphism_of_clifford} whose fiber is a finite model of composition.

Somehow in this language we encode the fact that the cut is linear in any extra degrees of freedom you add in. This can also be seen in the nature of what happens when you interact the $\mathscr{H}\!om$ bundle with the universal property defining $\cat{M}$.
